---
title: "1006-create-tree"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Tree Making

## First, convert plink output to a vcf and prune

```{sh, eval=FALSE}
#in outputs/1000
plink --tped 175-plink.tped --tfam 175-plink.tfam  --out binary --recode --allow-extra-chr --noweb
plink --ped binary.ped --map binary.map --recode vcf --allow-extra-chr -out recode
bcftools +prune -l 0.9 -w 10000 recode.vcf  -Ov -o recode.prune.vcf
```


Should reheader it so we know what the samples are....

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
load("outputs/103/pca-meta.rda")

bams <- read_tsv("bamlists/175.bamlist", col_names = FALSE) %>%
  mutate(vcf_name = str_replace_all(X1, pattern = "^data/|AN1711171/|BK1703222/|_sorted.*$", "")) %>%
  mutate(index = 1:n())
ad_meta <- left_join(bams, pca_meta)
labels<-ad_meta %>% select(index, Site) %>% mutate(Label=paste0(Site,"-",index)) %>% select(Label)
write_tsv(labels,"outputs/1000/labels.txt", col_names = FALSE)
```

Reheader...

```{sh, eval=FALSE}
bcftools reheader --samples outputs/1000/labels.txt -o outputs/1000/recode.prune.reheadered.vcf outputs/1000/recode.prune.vcf
```

## Next, convert to phylip 
```{sh, eval=FALSE}
#in outputs/1000
~/github/mccloud-rrt/vcf2phylip.py -i recode.prune.reheadered.vcf
source activate py3; ~/github/mccloud-rrt/103-remove-invariant.py -p recode.prune.reheadered.min4.phy -o recode.prune.reheadered.min4.asc.phy; source deactivate;
```

## Infer ML tree
```{sh, eval=FALSE}
#still in outputs/1000
iqtree -s recode.prune.reheadered.min4.asc.phy -st DNA -m GTR+G4+F+ASC -bb 1000 -alrt 1000
```

Plot ML tree
```{r, warning=FALSE, message=FALSE}
library(ggtree)

tree<-read.tree(file="outputs/1000/recode.prune.reheadered.min4.asc.phy.contree")
df<-as_tibble(as.data.frame(cbind(tree$tip.label))) %>%
  rename(Tip = V1) %>%
  mutate(Site = gsub("-\\d+", "", Tip))

# get meta
load("outputs/103/pca-meta.rda")
bams <- read_tsv("bamlists/175.bamlist", col_names = FALSE) %>%
  mutate(vcf_name = str_replace_all(X1, pattern = "^data/|AN1711171/|BK1703222/|_sorted.*$", "")) %>%
  mutate(index = 1:n())
ad_meta <- left_join(bams, pca_meta)

# remove spaces
ad_meta <- ad_meta %>% mutate(Tip = paste0(Site,"-",index))

df<-left_join(df,ad_meta)

t1<-ggtree(tree, layout="equal_angle") 

#Reordering for plotting
df$Region<- factor(df$Region, levels=c("Southeast","Southcentral", "Southwest", "Interior", "Northwest",
                                                   "North","Midwest"))        

 


t1 %<+% df + geom_tippoint(aes(shape=Region, angle=angle, fill=Region), cex=4) +  scale_shape_manual(values=c(24,21,22,22,22,22,23)) +  geom_treescale(x=0, y=-.1)

ggsave("outputs/1000/iqtree.pdf")

ggtree(tree) %<+% df + geom_tippoint(aes(shape=Region, angle=angle, fill=Region), cex=4) +
                       geom_tiplab(aes(label=Site), align=TRUE, linetype = "dashed", linesize=0.3, size=3) +
                       scale_shape_manual(values=c(24,21,22,22,22,22,23)) + geom_nodelab2() +
                       xlim(0,2)+
                       geom_treescale()

ggsave("outputs/1000/iqtree.rectangular.pdf", width=11, height=22)

#dtree<-ggtree(tree, layout="daylight")
#save(dtree, file="outputs/1000/dtree.rda")
load("outputs/1000/dtree.rda")
d <- dtree$data
d <- d[!d$isTip,]
d$label <- as.numeric(d$label)
d <- d[d$label > 90,]

dt<-dtree %<+% df + geom_point(data=d, alpha=0.5, cex=3) +
  geom_tippoint(aes(shape=Region, angle=angle, fill=Region), cex=4) +
  scale_shape_manual(values=c(24,21,21,22,22,22,23)) 

dt

ggsave("outputs/1000/iqtree.daylight.pdf", width=7, height=5.5)
```

#Network

First, generate a bootstrapped distance tree.
```{r, warning=FALSE, message=FALSE}
library(phangorn)
dat<-read.dna("outputs/1000/recode.prune.reheadered.min4.asc.phy", format="sequential")
dat<-as.phyDat(dat)
set.seed(1)
bs <- bootstrap.phyDat(dat, FUN = function(x)nj(dist.hamming(x)), 
    bs=100)
tree <- nj(dist.hamming(dat))
par("mar" = rep(1, 4))
tree <- plotBS(tree, bs, "unrooted", cex=0.5)

```

Then, use that tree to make a network. Improve aesthetics and alter consensus levels if desired (0.3 also is default) here.

```{r, warning=FALSE, message=FALSE}
cnet <- consensusNet(bs, .3)
edge.col <- createLabel(cnet, tree, "black", nomatch="red")

#Looks bad here as it is so big, so will duplicate into file
plot(cnet, show.edge.label = T, "2D", edge.color = edge.col,
                  col.edge.label = "blue", cex=.5)

pdf("outputs/1000/cnet.pdf", width=11, height=11)
plot(cnet, show.edge.label = T, "2D", edge.color = edge.col,
                  col.edge.label = "blue", cex=.5)
dev.off()
```

## Neighbor-Net

I can make one of these in splits tree, though needs to be agct coded and opened in splitstree. I don't want to try to plot it in r at the moment, though see https://cran.r-project.org/web/packages/phangorn/vignettes/IntertwiningTreesAndNetworks.html for how to to do that.   
```{r, eval=FALSE}
data<-read.dna("outputs/1000/recode.prune.reheadered.min4.asc.phy", format="sequential")
write.nexus.data(data, file="outputs/1000/recode.prune.reheadered.min4.asc.nex")

```

# SVDQuartets
Need a nice nexus file to append control block to. Control block previously "Region_index"
```{r, eval=FALSE}
names<-df %>% select(Region, index) %>% mutate(Names=paste0(Region,"_",index)) %>% select(Names)
renamed<-data
rownames(renamed)<-names$Names
write.nexus.data(renamed, file="outputs/1000/svdq/data.nex", interleaved = FALSE)
```

Create file for svdquartets, using 100000 quartets and 10 bootstraps for testing
```{sh, eval=FALSE}
#ouputs/1000/svdq
cat data.nex control.nex > input.nex
paup4a166_osx input.nex 
```

Plot.   

```{r, warning=FALSE, message=FALSE}
svd<-read.nexus(file="outputs/1000/svdq/svdMAF.tre")
ggtree(svd, layout="equal_angle")+geom_tiplab()+geom_nodelab2()
```

What about an exhaustive search with 100 bootstrap replicates?    
In progress.


