---
title: "1201-PCA-and-adxmiture"
author: "Mac Campbell"
date: "6/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)
library(RcppCNPy)
library(ggpubr)
library(ggrepel)
library(viridis)
library(scatterpie)
```

## PCA and Admixture
I'm going to double check for non-related individuals.    
See 1201.1-calculate-pca.sh      

Keeping 352 individuals after filtering (removing 6)
Read 352 samples and 27905 sites
Estimating population allele frequenciesEM (MAF) converged at iteration: 6
Number of sites after MAF filtering (0.05): 5682

Created two versions. 
Read 352 samples and 27905 sites

Estimating population allele frequencies
EM (MAF) converged at iteration: 6

Number of sites after MAF filtering (0.01): 8754

Download and view:

```{sh, eval=FALSE}
scp chinook:~/pike-time/outputs/1201/*.cov .
scp chinook:~/pike-time/outputs/1201/*admix.Q.npy .
scp chinook:~/pike-time/outputs/1201/*kinship* .
scp chinook:~/pike-time/outputs/1201/358-relate.unrelated.npy .
```

load meta    
```{r}
meta<-read_csv("metadata/358.csv") %>% mutate(Individual=1:n())
```


## Related individuals
```{r}
kin<-npyLoad("outputs/1201/358.kinship.npy")

#Convert upper half of matrix to data frame compatible 
kinrind<-which(upper.tri(kin, diag = TRUE), arr.ind = TRUE)
kin<-cbind(kinrind, kin[kinrind])

kindf<-as_tibble(kin) %>%
  rename(Ind1=row, Ind2=col, Kinship=V3) %>%
  filter(Ind1 != Ind2) %>%
  mutate(Mean=mean(Kinship))

#Can I replace Ind1 and Ind2 with individual labels???
df<-kindf %>% left_join(meta, by=c("Ind1"="Individual")) %>%
  left_join(meta, by=c("Ind2"="Individual")) 

#Ahh, and what are our "related" individuals
df %>% filter(Kinship>0.0625) %>% group_by(Site.x, Site.y) %>% summarize(Count=n())
```

```{r}
popdf<-df %>% filter(Site.x==Site.y) %>% mutate(Site=Site.x)

cols<-viridis(6)

ggplot(popdf)+geom_histogram(aes(x=Kinship), bins=50, fill="grey", color="grey") +
  geom_vline(xintercept =.25, col=cols[1]) + #Parent offspring/full sibs
  geom_vline(xintercept =.125, col=cols[2]) + #2nd degree
  geom_vline(xintercept =.0625, col=cols[3]) + # 3rd
  geom_vline(xintercept = .0313, col=cols[4]) +
  geom_vline(xintercept = .0156, col=cols[5]) +
  geom_vline(xintercept = 0.0, col=cols[6]) +
  scale_x_continuous(breaks=c(0,0.0156, 0.0313, 0.0625, 0.125, 0.25),
                     labels=c("0 / Unrelated","5th Deg.","4th Deg.","3rd Deg.", "2nd Deg.", "1st Deg.")) +
  theme_bw() +
  theme(axis.text.x = element_text(hjust=1, angle=45)) +
  theme(panel.grid=element_blank()) +
  ylab("Count") +
  xlab("Kinship Coefficient")+
  facet_wrap(Site ~ ., ncol=4, scales="free_y")

ggsave("outputs/1201/relatedness.jpg", width=16, height=14)
```
Let's look at the covariance matrix, must drop out related inds.

```{r}
vector<-npyLoad("outputs/1201/358-relate.unrelated.npy") %>% as.vector() %>% as_tibble()
meta$Kept<-vector
kept<-meta %>% filter(Kept != 0)

cov<-read_delim("outputs/1201/358-relate.cov", col_names=FALSE, delim=" ") %>% as.matrix()

```


```{r}
#' @param samples character vector with the individuals IDs in the order in which
#' they were passed in the bamlist to angsd.
#' @param cov covariance matrix
covar2pcs <- function(samples, cov) {
  
  
  eig <- eigen(cov, symm = TRUE)
  PC <- as.data.frame(eig$vectors) %>%
    as_tibble() %>%
    setNames(sprintf("PC-%02d", 1:ncol(.)))
  
  samtib <- tibble(sample = samples)
  
  list(
    PCs = bind_cols(samtib, PC),
    eigevalues = eig$values
  )
}
```


```{r}
pca <- covar2pcs(kept$`RAD identifier`, cov)

pca_long <- pca$PCs %>%
  tidyr::gather(., key = "PC", "val", -sample)

# then expand a grid of the possible comparisons (ordered)
expg <- expand.grid(sample = pca$PCs$sample,
                    PCx = sprintf("PC-%02d", 1:6),
                    PCy = sprintf("PC-%02d", 1:6),
                    stringsAsFactors = FALSE) %>%
  tibble::as_tibble()

# then left join the pca results onto that
pca_pairs <- dplyr::left_join(expg, pca_long, by = c("sample", "PCx" = "PC")) %>%
  dplyr::rename(val_x = val) %>%
  dplyr::left_join(pca_long, by = c("sample", "PCy" = "PC")) %>%
  dplyr::rename(val_y = val)

pp_meta <- pca_pairs %>%   # just keep the first 6 PCs around
  left_join(., kept, by = c("sample" = "RAD identifier")) %>%
  mutate(group = Site) 
```

Plot:    
```{r}
npc <- 4
pp_meta2 <- pp_meta %>%
  filter( (PCx %in% sprintf("PC-%02d", 1:npc)) & 
            (PCy %in% sprintf("PC-%02d", 1:npc)) )

eig <- eigen(cov, symm = TRUE)
var<-eig$values/sum(eig$values)
cumvar<-cumsum(eig$values)/sum(eig$values)

head(var)
head(cumvar)
```


```{r}
ggplot(pp_meta2, aes(x = val_x, y = val_y, color=Region)) +
  geom_point() +
  facet_grid(PCx ~ PCy)
```

```{r}
sub12<-pp_meta2 %>% filter( (PCx =="PC-01") & (PCy =="PC-02") )
centers12<-sub12 %>% group_by(Site) %>% mutate(MeanX=mean(val_x), MeanY=mean(val_y)) %>% select(Site, MeanX, MeanY) %>% unique()


pc12<-ggplot(sub12, aes(x = val_x, y = val_y, color=Site)) +
  geom_point(size = 2, alpha=0.75) +
  geom_text_repel(data=centers12, aes(x=MeanX, y=MeanY, label=Site), color="black", fontface='bold', size=2)+
  scale_fill_discrete(na.value = "white") + 
  theme_bw()+
  theme(panel.grid=element_blank())+
  xlab(paste("PC1", " ", round((100*var[1]),2), "%", sep = ""))+
  ylab(paste("PC2", " ", round((100*var[2]),2), "%", sep = ""))+
  theme(legend.position = "")+
  ggtitle("Related Individuals Excluded PCs 1 & 2")+
  theme(plot.title = element_text(hjust = 0.5))



sub13<-pp_meta2 %>% filter( (PCx =="PC-01") & (PCy =="PC-03") )
centers13<-sub13 %>% group_by(Site) %>% mutate(MeanX=mean(val_x), MeanY=mean(val_y)) %>% select(Site, MeanX, MeanY) %>% unique()

pc13<-ggplot(sub13, aes(x = val_x, y = val_y, color = Site)) +
  geom_point(size = 2, alpha=0.75) +
  geom_text_repel(data=centers13, aes(x=MeanX, y=MeanY, label=Site), color="black", fontface='bold', size=2)+
  scale_fill_discrete(na.value = "white") + 
  theme_bw()+
  theme(panel.grid=element_blank())+
  xlab(paste("PC1", " ", round((100*var[1]),2), "%", sep = ""))+
  ylab(paste("PC3", " ", round((100*var[3]),2), "%", sep = ""))+
  theme(legend.position = "")+
  ggtitle("Related Individuals Excluded PCs 1 & 3")+
  theme(plot.title = element_text(hjust = 0.5))



ggarrange(pc12, pc13, ncol = 2)
```

Redo with different covariance matrix

```{r}
cov<-read_delim("outputs/1201/358-relate-minmaf01.cov", col_names=FALSE, delim=" ") %>% as.matrix()

```

```{r}
pca <- covar2pcs(kept$`RAD identifier`, cov)

pca_long <- pca$PCs %>%
  tidyr::gather(., key = "PC", "val", -sample)

# then expand a grid of the possible comparisons (ordered)
expg <- expand.grid(sample = pca$PCs$sample,
                    PCx = sprintf("PC-%02d", 1:6),
                    PCy = sprintf("PC-%02d", 1:6),
                    stringsAsFactors = FALSE) %>%
  tibble::as_tibble()

# then left join the pca results onto that
pca_pairs <- dplyr::left_join(expg, pca_long, by = c("sample", "PCx" = "PC")) %>%
  dplyr::rename(val_x = val) %>%
  dplyr::left_join(pca_long, by = c("sample", "PCy" = "PC")) %>%
  dplyr::rename(val_y = val)

pp_meta <- pca_pairs %>%   # just keep the first 6 PCs around
  left_join(., kept, by = c("sample" = "RAD identifier")) %>%
  mutate(group = Site) 
```



Plot:    
```{r}
npc <- 4
pp_meta2 <- pp_meta %>%
  filter( (PCx %in% sprintf("PC-%02d", 1:npc)) & 
            (PCy %in% sprintf("PC-%02d", 1:npc)) )

eig <- eigen(cov, symm = TRUE)
var<-eig$values/sum(eig$values)
cumvar<-cumsum(eig$values)/sum(eig$values)

head(var)
head(cumvar)
```



```{r}
ggplot(pp_meta2, aes(x = val_x, y = val_y, color=Region)) +
  geom_point() +
  facet_grid(PCx ~ PCy)
```

```{r}
sub12<-pp_meta2 %>% filter( (PCx =="PC-01") & (PCy =="PC-02") )
centers12<-sub12 %>% group_by(Site) %>% mutate(MeanX=mean(val_x), MeanY=mean(val_y)) %>% select(Site, MeanX, MeanY) %>% unique()


pc12<-ggplot(sub12, aes(x = val_x, y = val_y, color=Site)) +
  geom_point(size = 2, alpha=0.75) +
  geom_text_repel(data=centers12, aes(x=MeanX, y=MeanY, label=Site), color="black", fontface='bold', size=2)+
  scale_fill_discrete(na.value = "white") + 
  theme_bw()+
  theme(panel.grid=element_blank())+
  xlab(paste("PC1", " ", round((100*var[1]),2), "%", sep = ""))+
  ylab(paste("PC2", " ", round((100*var[2]),2), "%", sep = ""))+
  theme(legend.position = "")+
  ggtitle("Related Individuals Excluded PCs 1 & 2")+
  theme(plot.title = element_text(hjust = 0.5))



sub13<-pp_meta2 %>% filter( (PCx =="PC-01") & (PCy =="PC-03") )
centers13<-sub13 %>% group_by(Site) %>% mutate(MeanX=mean(val_x), MeanY=mean(val_y)) %>% select(Site, MeanX, MeanY) %>% unique()

pc13<-ggplot(sub13, aes(x = val_x, y = val_y, color = Site)) +
  geom_point(size = 2, alpha=0.75) +
  geom_text_repel(data=centers13, aes(x=MeanX, y=MeanY, label=Site), color="black", fontface='bold', size=2)+
  scale_fill_discrete(na.value = "white") + 
  theme_bw()+
  theme(panel.grid=element_blank())+
  xlab(paste("PC1", " ", round((100*var[1]),2), "%", sep = ""))+
  ylab(paste("PC3", " ", round((100*var[3]),2), "%", sep = ""))+
  theme(legend.position = "")+
  ggtitle("Related Individuals Excluded PCs 1 & 3")+
  theme(plot.title = element_text(hjust = 0.5))



ggarrange(pc12, pc13, ncol = 2)
```

Structure like plot of admixture:
```{r}
k16<-npyLoad("outputs/1201/358-relate.admix.Q.npy") %>% as_tibble() 
#Need to rename all colnames smartly!
colnames(k16)<-gsub("V","Q",colnames(k16))
k16$Individual<-kept$Individual
k16$Site<-kept$Site
k16$Region<-kept$Region
k16 <- k16 %>% arrange(Site)
k16<-mutate(k16, Index=1:n())
```


```{r}
q16<-k16 %>% dplyr::select(Index, Site, Region, Q1, Q2, Q3, Q4, Q5, Q6, Q7, Q8, Q9, Q10, Q11, Q12, Q13, Q14, Q15, Q16) %>% gather(key=Ancestry, value=Q, 4:19)

pops2<-q16  %>% group_by(Site) %>% mutate(Start=min(Index), Stop=max(Index)) %>% 
  select(Site,Start,Stop) %>% unique() %>% 
  mutate(Position=round((Start+Stop)/2))
```


```{r}
kplot<-ggplot(q16) +
  geom_col(aes(x=Index, y=Q, fill=Ancestry), color="NA", size = 0, width = 1)+
  geom_segment(data=pops2, x = pops2$Start - 0.5, y=0, xend = pops2$Start-0.5, yend=1, alpha=0.9, size=0.25) +
  geom_segment(data=pops2, x = pops2$Stop[length(pops2$Stop)]  + 0.5, y=0, xend= pops2$Stop[length(pops2$Stop)] + 0.5, yend=1,  alpha=0.9,
               size=0.25) +
  geom_segment(x=0, xend= pops2$Stop[length(pops2$Stop)], y=1, yend=1, alpha=0.9, size=0.25) +
  theme(axis.text.x = element_text(angle=45, hjust=1)) +
  ylim(0,1.01) +
  xlim(-0.1, pops2$Stop[length(pops2$Stop)]+1) +
  theme(panel.background = element_blank())+
  scale_x_continuous(breaks=pops2$Position, labels=pops2$Site) +
  xlab("") +
    scale_fill_viridis_d(option="magma")
 # theme(axis.text.x = element_blank(), axis.ticks.x=element_blank()) +
#  theme(legend.position = "NA") +

kplot

ggsave("outputs/1201/pike-admix-k16.jpeg", width=11, height=8.5)
```

```{r}
kplotsc<-ggplot(q16%>% filter(Region=="Southcentral")) +
  geom_col(aes(x=Index, y=Q, fill=Ancestry), color="NA", size = 0, width = 1)+
  geom_segment(data=pops2, x = pops2$Start - 0.5, y=0, xend = pops2$Start-0.5, yend=1, alpha=0.9, size=0.25) +
  geom_segment(data=pops2, x = pops2$Stop[length(pops2$Stop)]  + 0.5, y=0, xend= pops2$Stop[length(pops2$Stop)] + 0.5, yend=1,  alpha=0.9,
               size=0.25) +
  geom_segment(x=0, xend= pops2$Stop[length(pops2$Stop)], y=1, yend=1, alpha=0.9, size=0.25) +
  theme(axis.text.x = element_text(angle=45, hjust=1)) +
  ylim(0,1.01) +
  xlim(-0.1, pops2$Stop[length(pops2$Stop)]+1) +
  theme(panel.background = element_blank())+
  scale_x_continuous(breaks=pops2$Position, labels=pops2$Site) +
  xlab("") +
  scale_fill_viridis_d(option="magma")
 # theme(axis.text.x = element_blank(), axis.ticks.x=element_blank()) +
#  theme(legend.position = "NA") +

kplotsc
ggsave("outputs/1201/pike-admix-k16-southcentral.jpeg", width=11, height=8.5)

```

Finally, I save a reduced bamlist

```{r}
reduced<-kept %>% select(String)
write_tsv(reduced, "bamlists/352.bamlist", col_names = FALSE)
```

##NGS Admix

Set this up like so:
`$HOME/angsd/misc/NGSadmix -likes outputs/1201/352.beagle.gz -K 2 -o outputs/1201/352-ngsadmix-k2 -P 3`

I have hardcoded this, maybe I can figure it out a smarter way.      
1. Need meta (kept object)
2. Need qopt.


```{r}
library(raster)
library(ggsn)
PlotAdmix<- function(file) {
  
q<-read_delim(file, delim=" ", col_names=FALSE)
#Make generic colnames
nums<-1:length(colnames(q))
mynames<-paste0("Q",nums)

qs<-length(colnames(q))-1

colnames(q)<-mynames

#Last col empty
q<-q[1:length(colnames(q))-1]

#Arrange Sites

r<-bind_cols(kept,q) %>% arrange(Region, Site) %>% mutate(Index=1:n())

r$Site<-factor(r$Site, levels=unique(r$Site))

rdf<-r %>% dplyr::select(Region, Site, Index,  colnames(q) ) %>% gather(key=Ancestry, value=Q, 4:(4+length(colnames(q))-1))

#Make names for structure-like plot
labels<-rdf  %>% group_by(Site) %>% mutate(Start=min(Index), Stop=max(Index)) %>% 
  dplyr::select(Site,Start,Stop) %>% unique() %>% 
  mutate(Position=round((Start+Stop)/2))

#Plot
ggplot(rdf) + 
  geom_col(aes(x=Index,y=Q, fill=Ancestry), color="NA", size = 0, width = 1)+
  geom_segment(data=labels, x = labels$Start - 0.5, y=0, xend = labels$Start-0.5, yend=1, alpha=0.9, size=0.25) +
  geom_segment(data=labels, x = labels$Stop[length(labels$Stop)]  + 0.5, y=0, xend=labels$Stop[length(labels$Stop)] + 0.5, yend=1,  alpha=0.9,
               size=0.25) +
  geom_segment(x=0, xend=labels$Stop[length(labels$Stop)], y=1, yend=1, alpha=0.9, size=0.25) +
  geom_segment(x=0, xend=labels$Stop[length(labels$Stop)], y=0, yend=-0.0, alpha=0.9, size=0.25) +
  theme_bw() + 
  theme(panel.grid = element_blank()) +
  scale_fill_viridis_d(option="magma") +
  ylim(-0.001,1.01) +
  xlim(-0.1, labels$Stop[length(labels$Stop)]+1) +
  scale_x_continuous(breaks=labels$Position, labels=labels$Site) +
  theme(axis.text.x = element_text(angle=45, hjust=1)) +
  xlab("") 

ggsave(paste0("outputs/1201/ngsd-admix-",qs,".jpg"), width=12, height=8)

#Plot introduced pops Q
intros<-r %>% dplyr::select(Region, Site, Latitude, Longitude, colnames(q)) %>% group_by(Region,Site) %>% summarise_all(list(mean)) %>% filter(Region=="Southcentral")

domain <- c(-152, -149, 60, 63)

# here eric implements something and I copy it
tidy_subset <- function(x, longlat) {
  x@data$id <- rownames(x@data)
  x.f <- broom::tidy(x) %>%
    dplyr::left_join(., x@data, by = "id") %>%
    dplyr::tbl_df() %>%
    filter(long > longlat[1],
           long < longlat[2],
           lat > longlat[3],
           lat < longlat[4])
}

nat.earth<-stack("~/github/mccloud-rrt/outputs/1000/NE2_HR_LC_SR_W_DR/NE2_HR_LC_SR_W_DR.tif")

nat.crop <- crop(nat.earth, y=extent(domain))

rast.table <- data.frame(xyFromCell(nat.crop, 1:ncell(nat.crop)),
                         getValues(nat.crop/255))


rast.table$rgb <- with(rast.table, rgb(NE2_HR_LC_SR_W_DR.1,
                                       NE2_HR_LC_SR_W_DR.2,
                                       NE2_HR_LC_SR_W_DR.3,
                                       1))

pies<-ggplot(intros, aes(x=Longitude, y=Latitude)) +
  geom_raster(data = rast.table, mapping = aes(x = x, y = y), fill = rast.table$rgb, interpolate = TRUE) +
  geom_scatterpie(data=intros, 
                    aes(x=Longitude, y=Latitude, group=Site, r=.1), 
                           cols=colnames(q)) +
  geom_text_repel(data=intros, aes(x=Longitude, y=Latitude, label=Site), color="black", fontface="bold", size=3) +
    ggsn::scalebar(x.min = -152, x.max= -149, y.min=60, y.max=63,
                 dist = 50, dist_unit = "km", st.size=4, height=.05, transform=TRUE, model = 'WGS84',
                 st.color="black") +
    scale_fill_viridis_d(option="magma") +
    coord_fixed(ratio=1, xlim = c(-152, -149), ylim=c(60,63)) +
  theme_bw()+
  theme(panel.background = element_blank())+
  theme(panel.grid=element_blank())

pies              
ggsave(paste0("outputs/1201/ngsd-admix-soutcentral-nodepies",qs,".jpg"), width=6, height=6)
}
```

```{r}
#file<-"outputs/1201/352-ngsadmix-k3.qopt"
#PlotAdmix(file)
files<-list.files("outputs/1201/", pattern="qopt", full.names = TRUE)
lapply(files, PlotAdmix)
```

Likes.     
```{sh, eval=FALSE}
"yes" | rm outputs/1201/likes.txt;
"yes" | rm outputs/1201/filenames.txt;

ls outputs/1201/ | grep log >> outputs/1201/filenames.txt
ls outputs/1201/ | grep log | while read line; do grep like outputs/1201/$line >> outputs/1201/likes.txt; done;

```

```{r}
lfiles<-read_tsv("outputs/1201/filenames.txt", col_names=c("File"))
lfiles$K<-gsub("352-ngsadmix-k|.log","",lfiles$File)
likers<-read_tsv("outputs/1201/likes.txt", col_names=c("String")) %>% separate(String, into=c("One","Likes","Three","Four"), sep=" ") %>% dplyr::select(Likes)
likers$Likes<-gsub("like=","",likers$Likes)


ddf<-bind_cols(lfiles,likers)
ddf$K<-as.numeric(ddf$K)
ddf$Likes<-as.numeric(ddf$Likes)
ddf<-ddf %>% arrange(K) %>% mutate(DK=Likes-lag(Likes))

ggplot(ddf) +
  geom_point(aes(x=K, y=DK))
```

Geographic plot.

```{r}
file<-c("outputs/1201/352-ngsadmix-k8.qopt")
q<-read_delim(file, delim=" ", col_names=FALSE)
#Make generic colnames
nums<-1:length(colnames(q))
mynames<-paste0("Q",nums)

qs<-length(colnames(q))-1

colnames(q)<-mynames

#Last col empty
q<-q[1:length(colnames(q))-1]

#Arrange Sites

r<-bind_cols(kept,q) %>% arrange(Region, Site) %>% mutate(Index=1:n())

r$Site<-factor(r$Site, levels=unique(r$Site))

rdf<-r %>% dplyr::select(Region, Site, Index,  Latitude, Longitude, colnames(q) ) %>% gather(key=Ancestry, value=Q, 6:(6+length(colnames(q))-1))

```


### Pull out "pure" pike
We have q16, and can merge with meta based on Kept

```{r}
keepers<-kept %>% mutate(Index=1:n()) %>% dplyr::select(-Site,-Region)

puredf<-q16 %>% filter(Q>0.90) %>% left_join(keepers)
write_csv(puredf, "metadata/q16-90.csv")
summarypuredf<-puredf %>% group_by(Site, Ancestry) %>% summarize(Count=n())
write_csv(summarypuredf, "metadata/q16-90-summary.csv")

```
