---
title: "1400-phylogeny"
author: "Mac Campbell"
date: "August 10, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)
library(adegenet)
library(vcfR)
library(ggtree)
library(ape)
library(ggpubr)
library(gridExtra)
```

## Generate a list of all wild samples

```{r}
bamlist<-read_tsv("bamlists/352.bamlist", col_names=c("String"))
meta<-read_csv("metadata/358.csv")
meta<-left_join(bamlist, meta)
```

```{r}
wgs<-read_csv("metadata/wgs-only.csv")

sub<-meta %>% dplyr::select(String, Region, Site, `RAD identifier`) %>% mutate(DataType="RAD") %>%
  filter(Region !="Southcentral") %>%
  bind_rows(wgs)

write_tsv(dplyr::select(sub,String), "bamlists/wgs-phy.bamlist", col_names=FALSE)
write_tsv(dplyr::select(sub,`RAD identifier`), "outputs/1400/samples.tsv", col_names = FALSE)
```

```{r}
sub %>% group_by(Site, Region, DataType) %>% summarize(Count=n())
```

We have 139 samples.

Generating a plink file, see 1400.1-plink.sh.


```{sh, eval=FALSE}
plink --tped plink.tped --tfam plink.tfam  --out binary --recode --allow-extra-chr --noweb
plink --ped binary.ped --map binary.map --recode vcf --allow-extra-chr -out recode
bcftools +prune -l 0.9 -w 10000 recode.vcf  -Ov -o recode.prune.vcf
bcftools reheader --samples samples.tsv -o recode.prune.reheadered.vcf recode.prune.vcf
source activate py2; ~/github/mccloud-rrt/vcf2phylip.py -i recode.prune.reheadered.vcf; conda deactivate;
 ~/github/mccloud-rrt/103-remove-invariant.py -p recode.prune.reheadered.min4.phy -o recode.prune.reheadered.min4.asc.phy
```
 
## PCA


Read in recode.prune.reheadered.vcf

```{r}
vcf<-read.vcfR(file="outputs/1400/recode.prune.reheadered.vcf")
genind<-vcfR2genind(vcf)
genind@pop<-as.factor(paste0(sub$Site,"-",sub$DataType))
```

425 variants.    


```{r}
gen<-genind
X <- tab(gen, NA.method="mean")

pca1 <- dudi.pca(X,scannf=FALSE,scale=FALSE,nf=3)
plot(pca1$li)
```

```{r}
pcdf<-as_tibble(pca1$l1) %>% rename(PC1=RS1, PC2=RS2, PC3=RS3)
pcdf$Site<-pop(gen)
pcdf<-pcdf %>% separate(Site, into=c("Site","DataType"))
pcdf<-pcdf %>% mutate(Region = ifelse(Site=="AntlinRiver","Southeast",
                         ifelse(Site %in% c("MintoFlats","YukonFlats","Fairbanks"), "Interior",
                         ifelse(Site=="NorthSlope","North",
                         ifelse(Site=="EagleLake","Midwest",
                         ifelse(Site=="Selawik","Northwest",
                         ifelse(Site %in% c("LakeClark","LakeNerka"), "Southwest",
                         ifelse(Site %in% c("PalmerLake","YukonRiver"),"Western Canada","NA"))))))))

pcdf$Region<-factor(pcdf$Region, levels=c("Southeast", "Southwest", "Interior", "Northwest",
                                                   "North","Western Canada","Midwest"))        

```

```{r}
eigens<-as_tibble(pca1$eig) %>% rename(Eigen=value) %>% mutate(Porportion=Eigen/sum(Eigen)*100)

ggplot(pcdf)+geom_point(aes(x=PC1, y=PC1, fill=Region, shape=DataType), alpha=0.75, size=2)+
  theme_bw()+
  theme(panel.grid=element_blank()) +
  theme(axis.title = element_text(face="bold")) +
  theme(legend.title = element_text(face="bold")) +
  scale_shape_manual(values = c(21,24)) +
  guides(fill = guide_legend(override.aes=list(shape=22))) +
  theme(legend.position = "") +
  theme(plot.title = element_text(face="bold", size=16)) 

ggplot(pcdf)+geom_point(aes(x=PC2, y=PC2, fill=Site, shape=DataType), alpha=0.75, size=2)+
  theme_bw()+
  theme(panel.grid=element_blank()) +
  theme(axis.title = element_text(face="bold")) +
  theme(legend.title = element_text(face="bold")) +
  scale_shape_manual(values = c(21,24)) +
  guides(fill = guide_legend(override.aes=list(shape=22))) +
  theme(legend.position = "") +
  theme(plot.title = element_text(face="bold", size=16)) 

```
Southeast->Antlen River->#F8766D
Southcentral->All the Introduced Pops->#C49A00
Southwest->Lake Nerka, Lake Clark->#53B400
Interior->Yukon Flats, Minto Flats, Fairbanks->#00C094
Northwest->Selawik River 00B6EB (Not in 2a/b)
North Slope->  #A58AFF (not in 2a/b)
Midwest -> Eagle Lake -> #FB61D7 -> Not in 2a/2b
```{r}
p12<-ggplot(pcdf)+geom_point(aes(x=PC1, y=PC2, fill=Region, shape=DataType), alpha=0.75, size=2)+
  xlab(paste0("PC1 ", round(eigens$Porportion[1],2),"%")) +
  ylab(paste0("PC2 ", round(eigens$Porportion[2],2),"%")) +
  theme_bw()+
  theme(panel.grid=element_blank()) +
  theme(axis.title = element_text(face="bold")) +
  theme(legend.title = element_text(face="bold")) +
  scale_shape_manual(values = c(21,24)) +
  scale_fill_manual(values = c("#F8766D","#53B400","#00C094","#00B6EB","#A58AFF","yellow","#FB61D7")) +
  guides(fill = guide_legend(override.aes=list(shape=22))) +
  theme(legend.position = "") +
  ggtitle("A")+
  theme(plot.title = element_text(face="bold", size=16))
p12
```

Nice to see Fairbanks fish cluster with Fairbanks fish.     

```{r}
ggplot(pcdf)+geom_point(aes(x=PC3, y=PC3, fill=Site, shape=DataType), alpha=0.75, size=2)+
  theme_bw()+
  theme(panel.grid=element_blank()) +
  theme(axis.title = element_text(face="bold")) +
  theme(legend.title = element_text(face="bold")) +
  scale_shape_manual(values = c(21,24)) +
  guides(fill = guide_legend(override.aes=list(shape=22))) +
  theme(legend.position = "") +
  theme(plot.title = element_text(face="bold", size=16))

```

```{r}
p13<-ggplot(pcdf)+geom_point(aes(x=PC1, y=PC3, fill=Region, shape=DataType), alpha=0.75, size=2)+
  xlab(paste0("PC1 ", round(eigens$Porportion[1],2),"%")) +
  ylab(paste0("PC3 ", round(eigens$Porportion[3],2),"%")) +
  theme_bw()+
  theme(panel.grid=element_blank()) +
  theme(axis.title = element_text(face="bold")) +
  theme(legend.title = element_text(face="bold")) +
  scale_shape_manual(values = c(21,24)) +
  scale_fill_manual(values = c("#F8766D","#53B400","#00C094","#00B6EB","#A58AFF","yellow","#FB61D7")) +
  guides(fill = guide_legend(override.aes=list(shape=22)))  +
  theme(legend.position = "") +
  ggtitle("B")+
  theme(plot.title = element_text(face="bold", size=16))
p13
```


## Tree
```{r}
labelling<-sub %>% dplyr::select(Site, DataType, `RAD identifier`) %>% mutate(Label=paste0(Site,DataType)) %>% group_by(Label) %>% 
  mutate(Text=paste(`RAD identifier`, collapse = " ")) %>% dplyr::select(Label, Text) %>% unique()

counts<-sub %>% dplyr::select(Site, DataType, `RAD identifier`) %>% mutate(Label=paste0(Site,DataType)) %>% 
    group_by(Label, Site, DataType) %>% summarize(Count=n()) %>%
    mutate(Region = ifelse(Site=="AntlinRiver","Southeast",
                         ifelse(Site %in% c("MintoFlats","YukonFlats","Fairbanks"), "Interior",
                         ifelse(Site=="NorthSlope","North",
                         ifelse(Site=="EagleLake","Midwest",
                         ifelse(Site=="Selawik","Northwest",
                         ifelse(Site %in% c("LakeClark","LakeNerka"), "Southwest",
                         ifelse(Site %in% c("PalmerLake","YukonRiver"),"Western Canada","NA"))))))))


counts$Region<-factor(counts$Region, levels=c("Southeast", "Southwest", "Interior", "Northwest",
                                                   "North","Western Canada","Midwest"))        

write_tsv(labelling,"outputs/1400/labels-for-nexus.tsv", col_names=FALSE)


```


Plotting     
```{r}
tree<-read.nexus("outputs/1400/svdMAF.tre")
t<-ggtree(tree, branch.length = "none")
```

```{r}
u<-t %<+% counts +
  geom_tippoint(aes(fill=Region, shape=DataType), size=3) +
  geom_tiplab(aes(label=Site), offset=.5) +
  geom_nodelab(fontface="bold") +
  xlim(0,10) +
  scale_shape_manual(values=c(21,24), name="Data Type") + 
  scale_fill_manual(values = c("#F8766D","#53B400","#00C094","#00B6EB","#A58AFF","yellow","#FB61D7")) +
  guides(fill = guide_legend(override.aes=list(shape=22))) +
  guides(size = FALSE) +
  theme(legend.title = element_text(face="bold")) +
  ggtitle("C")+
  theme(plot.title = element_text(face="bold", size=16))

u
ggsave("outputs/1400/test-tree.jpg")
```

Combine:
```{r}
ggarrange(arrangeGrob(p12,p13, ncol=1), 
          u, widths = c(1, 1.25))
ggsave("outputs/1400/figure.jpg", width=8.5, height = 7)
ggsave("outputs/1400/figure.pdf", width=8.5, height = 7)

```



# Plink2

```{sh, eval=FALSE}
plink --tped plink2.tped --tfam plink2.tfam  --out binary2 --recode --allow-extra-chr --noweb
plink --ped binary2.ped --map binary2.map --recode vcf --allow-extra-chr -out recode2
bcftools +prune -l 0.9 -w 10000 recode2.vcf  -Ov -o recode2.prune.vcf
bcftools reheader --samples samples.tsv -o recode2.prune.reheadered.vcf recode2.prune.vcf
source activate py2; ~/github/mccloud-rrt/vcf2phylip.py -i recode2.prune.reheadered.vcf; conda deactivate;
 ~/github/mccloud-rrt/103-remove-invariant.py -p recode2.prune.reheadered.min4.phy -o recode2.prune.reheadered.min4.asc.phy
```

 1226 sites pruned to 656 to 430 with ASC.
 
 Dropping selawik
 
 Plotting     
```{r}
tree2<-read.nexus("outputs/1400/svdMAF2.tre")
t2<-ggtree(tree2, branch.length = "none")
```

```{r}
t2 %<+% counts +
  geom_tippoint(aes(size=Count, fill=Site), pch=21) +
  geom_tiplab(offset=.5) +
  geom_nodelab(fontface="bold") +
  xlim(0,10)

ggsave("outputs/1400/test-tree-430-snps.jpg")
```
 