---
title: "1601-make-genepop"
author: "Mac Campbell"
date: "2022-10-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)
library(snpR)
```

Need colnames.

```{r}
m358<-read_tsv("~/Dropbox/ak-pike-genetics/EVA submission/supplemental/Supplemental Table S1.txt")
m358
```
```{r}
genos<-read_tsv(file="outputs/1601/358-dapc.geno", col_names=FALSE) %>% select(-X361)

colnames(genos)<-c("Chrom","Site",m358$`RAD identifier`)
genos
```

#make data long

```{r}
g1<-genos %>% mutate(Locus=paste0(Chrom,"-",Site)) %>% select(-Chrom,-Site) %>% relocate(Locus) #%>% rownames_to_column(var = "Locus")
rownames(g1)<-g1$Locus
g2<-g1 %>% select(-Locus)
g3<-t(g2)
g4<-as_tibble(g3)
colnames(g4)<-g1$Locus
g4$Pop<-paste0(m358$Site,",")
g4<-g4 %>% relocate(Pop)

g4
```

df[df == "Old Value"] <- "New Value"

```{r}
g5<-g4 %>% select(-Pop)
g5<-bind_cols(lapply(g5, function(x) gsub("A","111",x)))
g5<-bind_cols(lapply(g5, function(x) gsub("G","222",x)))
g5<-bind_cols(lapply(g5, function(x) gsub("C","333",x)))
g5<-bind_cols(lapply(g5, function(x) gsub("T","444",x)))
g5<-bind_cols(lapply(g5, function(x) gsub("N","000",x)))
g5$Pop<-g4$Pop
g5<-g5 %>% relocate(Pop)
g5<-g5 %>% arrange(Pop)
```


```{r}
write_delim(g5, delim=" ", file="outputs/1601/358-genepop.txt")
```

## Now to run it via snpR

```{sh, eval=FALSE}
plink --tped 358-dapc.tped --tfam 358-dapc.tfam  --out binary --recode --allow-extra-chr --noweb
plink --ped binary.ped --map binary.map --recode vcf --allow-extra-chr -out recode
bcftools +prune -l 0.9 -w 10000 recode.vcf  -Ov -o recode.prune.vcf
bcftools reheader --samples ../../metadata/358-samples.tsv -o recode.prune.reheadered.vcf recode.prune.vcf
```

now 2933 snps
```{r}
my.dat <- import.snpR.data("outputs/1601/recode.prune.reheadered.vcf", sample.meta = m358 %>% select(Site))
```

```{r}
ne<-calc_ne(
  my.dat,
  facets = "Site",
  chr = "CHROM",
  NeEstimator_path = "/Users/mac/Downloads/NeEstimator/Ne2-1M",
  mating = "random",
  pcrit = c(0.05, 0.02, 0.01,0),
  methods = c("LD"),
  max_ind_per_pop = NULL,
  outfile = "ne_out",
  verbose = TRUE,
  cleanup = TRUE
)
```

```{r}
res<-get.snpR.stats(ne, "Site", stat="ne")
write_tsv(res$pop, "outputs/1601/LD-method.tsv")
```

```{r}
#ne01<-res$pop %>% select(pop,"He_Ne_0.01","He_lCIp_0.01","He_uCIp_0.01")
#write_tsv(ne01, "outputs/1601/ne01.tsv")
```