---
title: "1800-treemix"
author: "Mac Campbell"
date: "2022-10-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)
```

## Treemix     
in `outputs/1800/`    

__1__ Data

```{sh, eval=FALSE}
ln -s ~/Dropbox/ak-pike-genetics/data-files/358-phylogeny/recode.prune.reheadered.vcf .
````
__1__ A text file with no header line, and one line per sample with the following three fields:

1 Family ID 2 Within-family ID 3 Cluster name

Example:     
CoyS1_2 CoyS1_2	Coy    
CoyS2_2 CoyS2_2	Coy    
CoyS3_2 CoyS3_2	Coy     
CoyS4_2 CoyS4_2	Coy    

```{sh, eval=FALSE}
cut -f 6 ~/Dropbox/ak-pike-genetics/EVA\ submission/supplemental/Supplemental\ Table\ S1.txt | awk '{print $1, $1}' | perl -pe 's/RAD.+\n//g' > 358-part1.txt

cut -f 9 ~/Dropbox/ak-pike-genetics/EVA\ submission/supplemental/Supplemental\ Table\ S1.txt | perl -pe 's/Sit.+\n//g' > 358-part2.txt

paste 358-part1.txt 358-part2.txt  > 358.clst
```

There may be a problem with chroms. Trying initially with recode.prune.reheadered.vcf 

Had to increase ulimit -n 3000     

```{sh, eval=FALSE}
bcftools view -H recode.prune.reheadered.vcf | cut -f 1 | uniq | awk '{print $0"\t"$0}' > redoneChroms.chrom-map.txt
vcftools --vcf  recode.prune.reheadered.vcf --plink --mac 2 --remove-indels --max-alleles 2  --out redoneChroms --chrom-map redoneChroms.chrom-map.txt

plink --file redoneChroms --make-bed --out redoneChroms --allow-no-sex --allow-extra-chr # drop 0
plink --bfile redoneChroms --freq --missing  --within 358.clst --out redoneChroms --allow-no-sex --allow-extra-chr

gzip redoneChroms.frq.strat
plink2treemix.py redoneChroms.frq.strat.gz redoneChroms.treemix.frq.gz
gunzip -c redoneChroms.treemix.frq.gz 
```

-noss ? -k appears to account for linkage. we pruned the data.
```{sh, eval=FALSE}
treemix -i redoneChroms.treemix.frq.gz -m 0 -o redone-noss-0 -root  EagleLake -global  -k 50  -noss > treemix_0_log
treemix -i redoneChroms.treemix.frq.gz -m 1 -o redone-noss-1 -root  EagleLake -global  -k 50  -noss > treemix_1_log
treemix -i redoneChroms.treemix.frq.gz -m 2 -o redone-noss-2 -root  EagleLake -global  -k 50  -noss > treemix_2_log
treemix -i redoneChroms.treemix.frq.gz -m 3 -o redone-noss-3 -root  EagleLake -global  -k 50  -noss > treemix_3_log
treemix -i redoneChroms.treemix.frq.gz -m 4 -o redone-noss-4 -root  EagleLake -global  -k 50  -noss > treemix_4_log

```

Plot    

```{r}
library(RColorBrewer)
library(R.utils)
source("~/treemix/treemix-1.13/src/plotting_funcs.R")
```

```{r}
prefix="outputs/1800/redone-noss-"

par(mfrow=c(3,2))
for(edge in 0:4){
  plot_tree(cex=0.8,paste0(prefix,edge))
  title(paste(edge,"edges"))
}
```

```{r}
pdf("outputs/1800/358-noss.pdf", height=12, width=8)

prefix="outputs/1800/redone-noss-"

par(mfrow=c(3,2))
for(edge in 0:4){
  plot_tree(cex=0.8,paste0(prefix,edge))
  title(paste(edge,"edges"))
}

dev.off()
```


###using sample size correction as we do have multiple samples/location

```{sh, eval=FALSE}
treemix -i redoneChroms.treemix.frq.gz -m 0 -o redone-0 -root  EagleLake -global  -k 50 > treemix_0_stand_log
treemix -i redoneChroms.treemix.frq.gz -m 1 -o redone-1 -root  EagleLake -global  -k 50 > treemix_1_stand_log
treemix -i redoneChroms.treemix.frq.gz -m 2 -o redone-2 -root  EagleLake -global  -k 50 > treemix_2_stand_log
treemix -i redoneChroms.treemix.frq.gz -m 3 -o redone-3 -root  EagleLake -global  -k 50 > treemix_3_stand_log
treemix -i redoneChroms.treemix.frq.gz -m 4 -o redone-4 -root  EagleLake -global  -k 50 > treemix_4_stand_log

```

```{r}
pdf("outputs/1800/358.pdf", height=12, width=8)

prefix="outputs/1800/redone-"

par(mfrow=c(3,2))
for(edge in 0:4){
  plot_tree(cex=0.8,paste0(prefix,edge))
  title(paste(edge,"edges"))
}

dev.off()
```



# Removing small samples sizes.   

```{sh}
cat outputs/1800/358-part2.txt | sort | uniq -c | sort -k 1 -n
```

Dropping Selawik River and Northslope (using 1801).

Getting reduced sample size list and vcf.    

```{sh, eval=FALSE}
cat ../1800/358.clst  | grep -v "Selawik" | grep -v "NorthSlope" > 352.clst
cut -f 1 -d ' ' 352.clst > samples-352.txt
bcftools view -S samples-352.txt ../1800/recode.prune.reheadered.vcf > recode.prune.reheadered.vcf
```

Now for some copy and paste!     

```{sh, eval=FALSE}
bcftools view -H recode.prune.reheadered.vcf | cut -f 1 | uniq | awk '{print $0"\t"$0}' > redoneChroms.chrom-map.txt
ulimit -n 3000   
vcftools --vcf  recode.prune.reheadered.vcf --plink --mac 2 --remove-indels --max-alleles 2  --out redoneChroms --chrom-map redoneChroms.chrom-map.txt

plink --file redoneChroms --make-bed --out redoneChroms --allow-no-sex --allow-extra-chr 
plink --bfile redoneChroms --freq --missing  --within 352.clst --out redoneChroms --allow-no-sex --allow-extra-chr

gzip redoneChroms.frq.strat
plink2treemix.py redoneChroms.frq.strat.gz redoneChroms.treemix.frq.gz
gunzip -c redoneChroms.treemix.frq.gz 
```

After filtering, kept 2015 out of a possible 2015 Sites (same as 358)

Running:

```{sh, eval=FALSE}
treemix -i redoneChroms.treemix.frq.gz -m 0 -o redone-noss-0 -root  EagleLake -global  -k 50  -noss > treemix_0_log
treemix -i redoneChroms.treemix.frq.gz -m 1 -o redone-noss-1 -root  EagleLake -global  -k 50  -noss > treemix_1_log
treemix -i redoneChroms.treemix.frq.gz -m 2 -o redone-noss-2 -root  EagleLake -global  -k 50  -noss > treemix_2_log
treemix -i redoneChroms.treemix.frq.gz -m 3 -o redone-noss-3 -root  EagleLake -global  -k 50  -noss > treemix_3_log
treemix -i redoneChroms.treemix.frq.gz -m 4 -o redone-noss-4 -root  EagleLake -global  -k 50  -noss > treemix_4_log

```


```{sh, eval=FALSE}
treemix -i redoneChroms.treemix.frq.gz -m 0 -o redone-0 -root  EagleLake -global  -k 50  > treemix_0_stand_log
treemix -i redoneChroms.treemix.frq.gz -m 1 -o redone-1 -root  EagleLake -global  -k 50  > treemix_1_stand_log
treemix -i redoneChroms.treemix.frq.gz -m 2 -o redone-2 -root  EagleLake -global  -k 50  > treemix_2_stand_log
treemix -i redoneChroms.treemix.frq.gz -m 3 -o redone-3 -root  EagleLake -global  -k 50  > treemix_3_stand_log
treemix -i redoneChroms.treemix.frq.gz -m 4 -o redone-4 -root  EagleLake -global  -k 50  > treemix_4_stand_log

```


```{r}
pdf("outputs/1801/352.pdf", height=12, width=8)

prefix="outputs/1801/redone-"

par(mfrow=c(3,2))
for(edge in 0:4){
  plot_tree(cex=0.8,paste0(prefix,edge))
  title(paste(edge,"edges"))
}

dev.off()
```
   

```{r}
pdf("outputs/1801/352-noss.pdf", height=12, width=8)

prefix="outputs/1801/redone-noss-"

par(mfrow=c(3,2))
for(edge in 0:4){
  plot_tree(cex=0.8,paste0(prefix,edge))
  title(paste(edge,"edges"))
}

dev.off()
```