---
title: "800-ngsAdmix"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## NGSadmix k=2


NGSadmix -likes inputBeagleFile.gz -K 3 -o outFileName -P 10
```{sh, eval=FALSE}
/Users/mac/angsd/misc/NGSadmix -likes outputs/102/175cov.beagle.gz -K 2 -o outputs/800/175cov-k2 -P 3

```

The default includes:  
Filters: misTol=0.050000 minMaf=0.050000 minLrt=0.000000 minInd=0  
Input file has dim: nsites=21523 nind=175  
Input file has dim (AFTER filtering): nsites=5694 nind=175  

### What does our qopt look like?

```{R k2, warning=FALSE, message=FALSE}
library(tidyverse)
q2<-read_delim("outputs/800/175cov-k2.qopt", delim=" ", col_names=FALSE) %>% select(X1, X2) %>%
  rename(Q1=X1, Q2=X2)

#Meta
load("outputs/103/pca-meta.rda")
bams <- read_tsv("bamlists/175.bamlist", col_names = FALSE) %>%
  mutate(vcf_name = str_replace_all(X1, pattern = "^data/|AN1711171/|BK1703222/|_sorted.*$", "")) %>%
  mutate(index = 1:n())
ad_meta <- left_join(bams, pca_meta)

# remove spaces
ad_meta <- ad_meta %>% mutate(Name = paste(gsub(Region, pattern=" ",
                                                   replacement=""), index, sep="_"))
# Separate Kenai
ad_meta <- ad_meta %>% mutate(Population = ifelse(Site == "StormyLake", "Kenai",
                                           ifelse(Site == "TinyLake", "Kenai",
                                           ifelse(Site %in% c("Tyonek", "ShellLake", "AndersonLake",
                                                              "OtterLake", "BulchitnaLake",
                                                              "AlexanderLake",
                                                              "YentnaIndianCreek", "DeshkaRiver"), 
                                                              "AnchorageBowl",
                                            Site)
)))

ad_meta$Population<- factor(ad_meta$Population, levels=c("AntlinRiver","EagleLake",
                                                         "MintoFlats","YukonFlats",
                                                         "Fairbanks", "AnchorageBowl",
                                                         "Kenai","LakeClark","LakeNerka",
                                                         "NorthSlope","Selawik"))
Population<-ad_meta %>% select(index, Name, Population)
df2<-cbind(q2, Population)                                                   
```

Plot it!

```{r ggplotk2, warning=FALSE, message=FALSE}

q2s<-df2 %>% select(Population, Name, index, Q1, Q2) %>% gather(key=Ancestry, value=Q, 4:5)

ggplot(q2s)+geom_col(aes(x=Name,y=Q, fill=Ancestry))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  scale_fill_brewer(palette ="Spectral")+
  facet_wrap(. ~ Population, ncol=4, scales="free_x")

```

## NGSadmix k=3

NGSadmix -likes inputBeagleFile.gz -K 3 -o outFileName -P 10
```{sh, eval=FALSE}
/Users/mac/angsd/misc/NGSadmix -likes outputs/102/175cov.beagle.gz -K 3 -o outputs/800/175cov-k3 -P 3
```
```{R k3, warning=FALSE, message=FALSE}

q3<-read_delim("outputs/800/175cov-k3.qopt", delim=" ", col_names=FALSE) %>% 
  select(X1, X2, X3) %>%rename(Q1=X1, Q2=X2, Q3=X3)

df3<-cbind(q3, Population)     

q3s<-df3 %>% select(Population, Name, index, Q1, Q2, Q3) %>% gather(key=Ancestry, value=Q, 4:6)

ggplot(q3s)+geom_col(aes(x=Name,y=Q, fill=Ancestry))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  scale_fill_brewer(palette ="Spectral")+
  facet_wrap(. ~ Population, ncol=4, scales="free_x")
```


## NGSadmix k=4

```{sh, eval=FALSE}
/Users/mac/angsd/misc/NGSadmix -likes outputs/102/175cov.beagle.gz -K 4 -o outputs/800/175cov-k4 -P 3
```

```{R k4, warning=FALSE, message=FALSE}

q4<-read_delim("outputs/800/175cov-k4.qopt", delim=" ", col_names=FALSE) %>% 
  select(X1, X2, X3, X4) %>%rename(Q1=X1, Q2=X2, Q3=X3, Q4=X4)

df4<-cbind(q4, Population)     

q4s<-df4 %>% select(Population, Name, index, Q1, Q2, Q3, Q4) %>% 
  gather(key=Ancestry, value=Q, 4:7)

ggplot(q4s)+geom_col(aes(x=Name,y=Q, fill=Ancestry))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  scale_fill_brewer(palette ="Spectral")+
  facet_wrap(. ~ Population, ncol=4, scales="free_x")
```

## NGSadmix k=5

```{sh, eval=FALSE}
/Users/mac/angsd/misc/NGSadmix -likes outputs/102/175cov.beagle.gz -K 5 -o outputs/800/175cov-k5 -P 3
```

```{R k5, warning=FALSE, message=FALSE}

q5<-read_delim("outputs/800/175cov-k5.qopt", delim=" ", col_names=FALSE) %>% 
  select(X1, X2, X3, X4, X5) %>%rename(Q1=X1, Q2=X2, Q3=X3, Q4=X4, Q5=X5)

df5<-cbind(q5, Population)     

q5s<-df5 %>% select(Population, Name, index, Q1, Q2, Q3, Q4, Q5) %>% 
  gather(key=Ancestry, value=Q, 4:8)

ggplot(q5s)+geom_col(aes(x=Name,y=Q, fill=Ancestry))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  scale_fill_brewer(palette ="Spectral")+
  facet_wrap(. ~ Population, ncol=4, scales="free_x")
```

## NGSadmix k=6

```{sh, eval=FALSE}
/Users/mac/angsd/misc/NGSadmix -likes outputs/102/175cov.beagle.gz -K 6 -o outputs/800/175cov-k6 -P 3
```

```{R k6, warning=FALSE, message=FALSE}

q6<-read_delim("outputs/800/175cov-k6.qopt", delim=" ", col_names=FALSE) %>% 
  select(X1, X2, X3, X4, X5, X6) %>%rename(Q1=X1, Q2=X2, Q3=X3, Q4=X4, Q5=X5, Q6=X6)

df6<-cbind(q6, Population)     

q6s<-df6 %>% select(Population, Name, index, Q1, Q2, Q3, Q4, Q5, Q6) %>% 
  gather(key=Ancestry, value=Q, 4:9)

ggplot(q6s)+geom_col(aes(x=Name,y=Q, fill=Ancestry))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  scale_fill_brewer(palette ="Spectral")+
  facet_wrap(. ~ Population, ncol=4, scales="free_x")
```

## NGSadmix k=7

```{sh, eval=FALSE}
/Users/mac/angsd/misc/NGSadmix -likes outputs/102/175cov.beagle.gz -K 7 -o outputs/800/175cov-k7 -P 3
```

```{R k7, warning=FALSE, message=FALSE}

q7<-read_delim("outputs/800/175cov-k7.qopt", delim=" ", col_names=FALSE) %>% 
  select(X1, X2, X3, X4, X5, X6, X7) %>%rename(Q1=X1, Q2=X2, Q3=X3, Q4=X4, Q5=X5, Q6=X6, Q7=X7)

df7<-cbind(q7, Population)     

q7s<-df7 %>% select(Population, Name, index, Q1, Q2, Q3, Q4, Q5, Q6, Q7) %>% 
  gather(key=Ancestry, value=Q, 4:10)

ggplot(q7s)+geom_col(aes(x=Name,y=Q, fill=Ancestry))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  scale_fill_brewer(palette ="Spectral")+
  facet_wrap(. ~ Population, ncol=4, scales="free_x")
```

## NGSadmix k=8

```{sh, eval=FALSE}
/Users/mac/angsd/misc/NGSadmix -likes outputs/102/175cov.beagle.gz -K 8 -o outputs/800/175cov-k8 -P 3
```


```{R k8, warning=FALSE, message=FALSE}

q8<-read_delim("outputs/800/175cov-k8.qopt", delim=" ", col_names=FALSE) %>% 
  select(X1, X2, X3, X4, X5, X6, X7, X8) %>%rename(Q1=X1, Q2=X2, Q3=X3, Q4=X4, Q5=X5, Q6=X6, Q7=X7, Q8=X8)

df8<-cbind(q8, Population)     

q8s<-df8 %>% select(Population, Name, index, Q1, Q2, Q3, Q4, Q5, Q6, Q7, Q8) %>% 
  gather(key=Ancestry, value=Q, 4:11)

ggplot(q8s)+geom_col(aes(x=Name,y=Q, fill=Ancestry))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  scale_fill_brewer(palette ="Spectral")+
  facet_wrap(. ~ Population, ncol=4, scales="free_x")
```

## NGSadmix k=9

```{sh, eval=FALSE}
/Users/mac/angsd/misc/NGSadmix -likes outputs/102/175cov.beagle.gz -K 9 -o outputs/800/175cov-k9 -P 3
```


```{R k9, warning=FALSE, message=FALSE}

q9<-read_delim("outputs/800/175cov-k9.qopt", delim=" ", col_names=FALSE) %>% 
  select(X1, X2, X3, X4, X5, X6, X7, X8, X9) %>%rename(Q1=X1, Q2=X2, Q3=X3, Q4=X4, Q5=X5, Q6=X6, Q7=X7, Q8=X8, Q9=X9)

df9<-cbind(q9, Population)     

q9s<-df9 %>% select(Population, Name, index, Q1, Q2, Q3, Q4, Q5, Q6, Q7, Q8, Q9) %>% 
  gather(key=Ancestry, value=Q, 4:12)

ggplot(q9s)+geom_col(aes(x=Name,y=Q, fill=Ancestry))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  scale_fill_brewer(palette ="Spectral")+
  facet_wrap(. ~ Population, ncol=4, scales="free_x")
```