---
title: "801-ngsAdmix-subset"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Let's get a few fish together from a subset of pops:

"AntlinRiver"
"EagleLake"
"Fairbanks"
"MintoFlats"
"YukonFlats"
"LakeClark"
"LakeNerka"  

```{r bamlist, warning=FALSE, message=FALSE}
library("tidyverse")
#Meta
load("outputs/103/pca-meta.rda")
bams <- read_tsv("bamlists/175.bamlist", col_names = FALSE) %>%
  mutate(vcf_name = str_replace_all(X1, pattern = "^data/|AN1711171/|BK1703222/|_sorted.*$", "")) %>%
  mutate(index = 1:n())
ad_meta <- left_join(bams, pca_meta)

# remove spaces
ad_meta <- ad_meta %>% mutate(Name = paste(gsub(Region, pattern=" ",
                                                   replacement=""), index, sep="_"))
# Separate Kenai
ad_meta <- ad_meta %>% mutate(Population = ifelse(Site == "StormyLake", "Kenai",
                                           ifelse(Site == "TinyLake", "Kenai",
                                           ifelse(Site %in% c("Tyonek", "ShellLake", "AndersonLake",
                                                              "OtterLake", "BulchitnaLake",
                                                              "AlexanderLake",
                                                              "YentnaIndianCreek", "DeshkaRiver"), 
                                                              "AnchorageBowl",
                                            Site)
)))

ad_meta$Population<- factor(ad_meta$Population, levels=c("AntlinRiver","EagleLake",
                                                         "MintoFlats","YukonFlats",
                                                         "Fairbanks", "AnchorageBowl",
                                                         "Kenai","LakeClark","LakeNerka",
                                                         "NorthSlope","Selawik"))

sub_meta<-filter(ad_meta, Population %in% c("AntlinRiver","EagleLake","Fairbanks",
                                            "MintoFlats","YukonFlats","LakeClark",
                                            "LakeNerka"))
save(sub_meta, file="outputs/sub_meta.rda")

#Write a bamlist

subbams<-sub_meta %>% select(`Sequencing project`, `RAD identifier`, Region, Site) %>% separate(`Sequencing project`, c("GW", "Project")) %>% mutate(String=paste("data/",Project,"/",`RAD identifier`,"_sorted.bam", sep=""))

write(subbams$String, file="bamlists/133.txt")
write(subbams$String, file="bamlists/133.bamlist")

```

# Now a short aside with 801.1-generate-pca-133.sh
Done....  
outputs/800 now has beagle file and .cov/.qopt  
Let's look at the PCA! Does it check out?  


```{r, warning=FAlSE, message=FALSE}
#' @param samples character vector with the individuals IDs in the order in which
#' they were passed in the bamlist to angsd.
#' @param cov covariance matrix
covar2pcs <- function(samples, cov) {
  
  
  eig <- eigen(cov, symm = TRUE)
  PC <- as.data.frame(eig$vectors) %>%
    as_tibble() %>%
    setNames(sprintf("PC-%02d", 1:ncol(.)))
  
  samtib <- tibble(sample = samples)
  
  list(
    PCs = bind_cols(samtib, PC),
    eigevalues = eig$values
  )
}
```

Shameless copying from Eric A. again.
```{r, warning=FALSE, message=FALSE}
cov<-read_tsv("outputs/800/133.cov", col_names=FALSE) %>% as.matrix()

pca <- covar2pcs(sub_meta$vcf_name, cov)

pca_long <- pca$PCs %>%
  tidyr::gather(., key = "PC", "val", -sample)

# then expand a grid of the possible comparisons (ordered)
expg <- expand.grid(sample = pca$PCs$sample,
                    PCx = sprintf("PC-%02d", 1:6),
                    PCy = sprintf("PC-%02d", 1:6),
                    stringsAsFactors = FALSE) %>%
  tibble::as_tibble()

# then left join the pca results onto that
pca_pairs <- dplyr::left_join(expg, pca_long, by = c("sample", "PCx" = "PC")) %>%
  dplyr::rename(val_x = val) %>%
  dplyr::left_join(pca_long, by = c("sample", "PCy" = "PC")) %>%
  dplyr::rename(val_y = val)

pp_meta <- pca_pairs %>%   # just keep the first 6 PCs around
  left_join(., pca_meta, by = c("sample" = "vcf_name")) %>%
  mutate(group = Region)  # put a group column in there

# now, that has the first 6 PCs in it.  If we want to focus on the just the
# first 3, we could do 
npc <- 3
pp_meta2 <- pp_meta %>%
  filter( (PCx %in% sprintf("PC-%02d", 1:npc)) & 
            (PCy %in% sprintf("PC-%02d", 1:npc)) )

ggplot(pp_meta2, aes(x = val_x, y = val_y, fill = group)) +
  geom_point(pch = 21, size = 2) +
  scale_fill_discrete(na.value = "white") + 
  facet_grid(PCy ~ PCx, scales = "free")


```

#On to ngsAdmix!
Let's plot K = 2,3,4,5,6 while we are at it.  

```{sh, eval=FALSE}
/Users/mac/angsd/misc/NGSadmix -likes outputs/800/133.beagle.gz -K 2 -o outputs/800/133-k2 -P 3
/Users/mac/angsd/misc/NGSadmix -likes outputs/800/133.beagle.gz -K 3 -o outputs/800/133-k3 -P 3
/Users/mac/angsd/misc/NGSadmix -likes outputs/800/133.beagle.gz -K 4 -o outputs/800/133-k4 -P 3
/Users/mac/angsd/misc/NGSadmix -likes outputs/800/133.beagle.gz -K 5 -o outputs/800/133-k5 -P 3
/Users/mac/angsd/misc/NGSadmix -likes outputs/800/133.beagle.gz -K 6 -o outputs/800/133-k6 -P 3
```

Here we can split things into runs, or somehow combine it all.  I think I will generate separate ggplot objects and plot with ggarrange of the ggpubr library.  Should be beautiful.

```{r k2, warning=FALSE, message=FALSE}
#Get metadata together, looks like factor levels of Population were retained from sub_meta into df2 below.
Population<-sub_meta %>% select(index, Name, Population)

q2<-read_delim("outputs/800/133-k2.qopt", delim=" ", col_names=FALSE) %>% select(X1, X2) %>%
  rename(Q1=X1, Q2=X2)

df2<-cbind(q2, Population)                               

q2s<-df2 %>% select(Population, Name, index, Q1, Q2) %>% gather(key=Ancestry, value=Q, 4:5)

p2<-ggplot(q2s)+geom_col(aes(x=Name,y=Q, fill=Ancestry))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  scale_fill_brewer(palette ="Spectral")+
  facet_wrap(. ~ Population, ncol=7, scales="free_x")

p2
```

Can I generalize this to a function?  
Let's copy Eric's grammar, but not his function.  

```{r myplotfun, warning=FALSE, message=FALSE}
#' @param file what file to read in
#' @param meta data frame with metadata
#' returns a ggplot object for plotting
admixPlot <- function(file, meta) {
q <-read_delim(paste(file), delim=" ", col_names=FALSE) 
q<-q[1:(length(colnames(q)-1))]

#And I'm bored. Also, we'll need to adjust the colors manually in the end to get them looking logical anywho. 
}
```

```{r k3, message=FAlSE, warning=FALSE}

q3<-read_delim("outputs/800/133-k3.qopt", delim=" ", col_names=FALSE) %>% select(X1, X2, X3) %>%
  rename(Q1=X1, Q2=X2, Q3=X3)
q4<-read_delim("outputs/800/133-k4.qopt", delim=" ", col_names=FALSE) %>% select(X1, X2, X3, X4) %>%
  rename(Q1=X1, Q2=X2, Q3=X3, Q4=X4)
q5<-read_delim("outputs/800/133-k5.qopt", delim=" ", col_names=FALSE) %>% select(X1, X2, X3, X4, X5) %>%
  rename(Q1=X1, Q2=X2, Q3=X3, Q4=X4, Q5=X5)
q6<-read_delim("outputs/800/133-k6.qopt", delim=" ", col_names=FALSE) %>% select(X1, X2, X3, X4, X5, X6) %>%
  rename(Q1=X1, Q2=X2, Q3=X3, Q4=X4, Q5=X5, Q6=X6)

df3<-cbind(q3, Population)                               
df4<-cbind(q4, Population)
df5<-cbind(q5, Population)                               
df6<-cbind(q6, Population)


q3s<-df3 %>% select(Population, Name, index, Q1, Q2, Q3) %>% 
  gather(key=Ancestry, value=Q, 4:6)
q4s<-df4 %>% select(Population, Name, index, Q1, Q2, Q3, Q4) %>% 
  gather(key=Ancestry, value=Q, 4:7)
q5s<-df5 %>% select(Population, Name, index, Q1, Q2, Q3, Q4, Q5) %>% 
  gather(key=Ancestry, value=Q, 4:8)
q6s<-df6 %>% select(Population, Name, index, Q1, Q2, Q3, Q4, Q5, Q6) %>% 
  gather(key=Ancestry, value=Q, 4:9)

```

```{r plots, warning=FALSE, message=FALSE}
p3<-ggplot(q3s)+geom_col(aes(x=Name,y=Q, fill=Ancestry))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  scale_fill_brewer(palette ="Spectral")+
  facet_wrap(. ~ Population, ncol=7, scales="free_x")

p4<-ggplot(q4s)+geom_col(aes(x=Name,y=Q, fill=Ancestry))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  scale_fill_brewer(palette ="Spectral")+
  facet_wrap(. ~ Population, ncol=7, scales="free_x")

p5<-ggplot(q5s)+geom_col(aes(x=Name,y=Q, fill=Ancestry))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  scale_fill_brewer(palette ="Spectral")+
  facet_wrap(. ~ Population, ncol=7, scales="free_x")

p6<-ggplot(q6s)+geom_col(aes(x=Name,y=Q, fill=Ancestry))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  scale_fill_brewer(palette ="Spectral")+
  facet_wrap(. ~ Population, ncol=7, scales="free_x")
```

#Now for the one plot to bind them all.

```{r plotting, warning=FALSE, message=FALSE}
library(ggpubr)

ggarrange(p2,p3,p4,p5,p6, ncol=1)

ggsave("outputs/800/133-k2-k6.pdf", units="in", width=11, height=8.5)
```

#Now to recolor manually

```{r colors, warning=FALSE, message=FALSE}
library(RColorBrewer)
cols<-brewer.pal(6, "Spectral")
ggplot()+geom_point(aes(x=c(1,2,3,4,5,6), y=c(1,1,1,1,1,1)), col=cols, cex=5)
```

```{r p2, warning=FALSE}
x2<-p2+scale_fill_manual(values=c(cols[3], cols[1]))+theme(legend.position = "none")
x3<-p3+scale_fill_manual(values=c(cols[1], cols[3], cols[2]))+theme(legend.position = "none")
x4<-p4+scale_fill_manual(values=c(cols[3], cols[6], cols[1], cols[2]))+theme(legend.position = "none")
x5<-p5+scale_fill_manual(values=c(cols[1], cols[3], cols[4], cols[6], cols[2])) +theme(legend.position = "none")

ggarrange(x2,x3,x4,x5, ncol=1)

ggsave("outputs/800/133-k2-k5.pdf", units="in", width=11, height=6)
```
