---
title: "903-process-diversity"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Watterson's Theta and Tajima's D
Can calculate like so:  
`~/angsd/misc/thetaStat do_stat AnchorageBowl-per-site.thetas.idx`

Format is:

(indexStart,indexStop)(posStart,posStop)(regStat,regStop) chrname wincenter tW tP tF tH tL tajD fulif fuliD fayH zengsE numSites


```{r, warning=FALSE, message=FALSE}
library(tidyverse)
df<-read_tsv("outputs/902/AnchorageBowl-per-site.thetas.idx.pestPG")
```

Hmmm.. .This does not provide actual numbers for the chromos. Switiching to a sliding window appraoch.

`~/angsd/misc/thetaStat do_stat AnchorageBowl-per-site.thetas.idx -win 50000 -step 10000 -outnames theta.thetasWindow.gz`

```{r, warning=FALSE, message=FALSE}
df2<-read_tsv("outputs/902/theta.thetasWindow.gz.pestPG")
```

Looking good! Now for some weighted means based on the number of sites used to calculate stats in the window.

```{r, warning=FALSE, message=FALSE}
library(matrixStats)
df2<-df2 %>% select(-`#(indexStart,indexStop)(firstPos_withData,lastPos_withData)(WinStart,WinStop)`) %>%
  select(Chr, WinCenter, tW, tP, Tajima, nSites)
df3<-df2 %>% filter(tW != "NaN") %>% filter(tP != "NaN")

tWWeight<-weightedMedian(df3$tW, w=df3$nSites)
tPWeight<-weightedMedian(df3$tP, w=df3$nSites)
TajimaWeight<-weightedMedian(df3$Tajima, w=df3$nSites)
```

Plot
```{r, warning=FALSE, message=FALSE}
df4<-mutate(df3, Population="Anchorage Bowl")
ggplot(df4)+geom_boxplot(aes(x=Population, y=tW, weight=nSites),
                          outlier.size=0.1, outlier.alpha=0.5, outlier.shape=15,
                          alpha=0.5)+
  theme_classic()

ggplot(df4)+geom_boxplot(aes(x=Population, y=Tajima, weight=nSites),
                          outlier.size=0.1, outlier.alpha=0.5, outlier.shape=15,
                          alpha=0.5)+
  theme_classic()
```
