---
title: "903-process-diversity"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Watterson's Theta and Tajima's D
Can calculate like so:  
`~/angsd/misc/thetaStat do_stat AnchorageBowl-per-site.thetas.idx`

Format is:

(indexStart,indexStop)(posStart,posStop)(regStat,regStop) chrname wincenter tW tP tF tH tL tajD fulif fuliD fayH zengsE numSites


```{r, warning=FALSE, message=FALSE}
library(tidyverse)
df<-read_tsv("outputs/902/AnchorageBowl-per-site.thetas.idx.pestPG")
```

Hmmm.. .This does not provide actual numbers for the chromos. Switiching to a sliding window appraoch.

`~/angsd/misc/thetaStat do_stat AnchorageBowl-per-site.thetas.idx -win 50000 -step 10000 -outnames theta.thetasWindow.gz`

```{r, warning=FALSE, message=FALSE}
df2<-read_tsv("outputs/902/AnchorageBowl.thetas-windows.gz.pestPG")
```

Looking good! Now for some weighted means based on the number of sites used to calculate stats in the window.

```{r, warning=FALSE, message=FALSE}
library(matrixStats)
df2<-df2 %>% select(-`#(indexStart,indexStop)(firstPos_withData,lastPos_withData)(WinStart,WinStop)`) %>%
  select(Chr, WinCenter, tW, tP, Tajima, nSites)
df3<-df2 %>% filter(tW != "NaN") %>% filter(tP != "NaN")

tWWeight<-weightedMedian(df3$tW, w=df3$nSites)
tPWeight<-weightedMedian(df3$tP, w=df3$nSites)
TajimaWeight<-weightedMedian(df3$Tajima, w=df3$nSites)
```

Plot
```{r, warning=FALSE, message=FALSE}
df4<-mutate(df3, Population="Anchorage Bowl")
ggplot(df4)+geom_boxplot(aes(x=Population, y=tW, weight=nSites),
                          outlier.size=0.1, outlier.alpha=0.5, outlier.shape=15,
                          alpha=0.5)+
  theme_classic()

ggplot(df4)+geom_boxplot(aes(x=Population, y=Tajima, weight=nSites),
                          outlier.size=0.1, outlier.alpha=0.5, outlier.shape=15,
                          alpha=0.5)+
  theme_classic()
```

#Everybody!

```{r tidyall, warning=FALSE, message=FALSE}
#Copied from here: https://serialmentor.com/blog/2016/6/13/reading-and-combining-many-tidy-data-files-in-R

#files<-c("AntlinRiver.thetas-windows.gz.pestPG", "NorthSlope.thetas-windows.gz.pestPG")
data_path<-"outputs/902"
files<-dir(data_path, pattern="*.thetas-windows.gz.pestPG")

data <- data_frame(filename = files) %>% # create a data frame
                                         # holding the file names
  mutate(file_contents = map(filename,          # read files into
           ~ read_tsv(file.path(data_path, .))) # a new data column
        ) 

```

Now, let's filter "data" to have what we want.
.thetas-windows.gz.pestPG

```{r filter, warning=FALSE, message=FALSE}
data2<- unnest(data) %>% select(-`#(indexStart,indexStop)(firstPos_withData,lastPos_withData)(WinStart,WinStop)`) %>% mutate(Population=gsub(".thetas-windows.gz.pestPG", "", filename)) %>%
  select(Population, Chr, WinCenter, nSites, tW, tP, Tajima) %>% filter(tW != "NaN") %>% filter(tP != "NaN")

#Reorder population
data2$Population<- factor(data2$Population, levels=c("AntlinRiver","EagleLake",
                                                         "MintoFlats","YukonFlats",
                                                         "Fairbanks", "AnchorageBowl",
                                                         "Kenai","LakeClark","LakeNerka",
                                                         "NorthSlope","Selawik"))


#compute stats
data2 %>% group_by(Population) %>% mutate(medianTW=weightedMedian(tW, w = nSites)) %>% mutate(medianD=weightedMedian(Tajima, w=nSites)) %>% select(Population, medianTW, medianD) %>% 
  summarize(ThetaW=unique(medianTW), Tajimas=unique(medianD), Count=n())
#tidy up
tidy<-data2 %>% select(Population, Chr, WinCenter, nSites, tW, Tajima) %>% gather(Statistic, Value, 5:6)

```

Snazzy plot!!!
```{r snazzyplot, warning=FALSE, message=FALSE}
tidy$Statistic<-factor(tidy$Statistic, levels=c("tW","Tajima"))
ggplot(tidy)+geom_boxplot(aes(x=Population, y=Value, weight=nSites),
                          outlier.size=0.1, outlier.alpha=0.5, outlier.shape=15,
                          alpha=0.5)+
  theme_bw() +
  theme(panel.grid=element_blank())+
  facet_grid(Statistic ~ ., scales="free_y")
ggsave(filename = "outputs/903/diversity-stats.pdf", width=11, height=8.5)
```

Filter out the small pops.
```{r}
tidybig<-filter(tidy, ! (Population %in% c("EagleLake","NorthSlope","Selawik") ) )
                
ggplot(tidybig)+geom_boxplot(aes(x=Population, y=Value, weight=nSites),
                          outlier.size=0.1, outlier.alpha=0.5, outlier.shape=15,
                          alpha=0.5)+
  theme_bw() +
  theme(panel.grid=element_blank()) +
  theme(axis.text.x=element_text(size=12, face="bold", angle=45, hjust=1)) +
  theme(axis.text.y=element_text(size=12)) +
  facet_grid(Statistic ~ ., scales="free_y")
  
ggsave(filename = "outputs/903/diversity-stats-bigpops.pdf", width=11, height=8.5)

```
